---
title: 'STAT 344 Project'
output: pdf_document
---
```{r}
library(tidyverse)
dat <- read.csv('crimedata_csv_AllNeighbourhoods_2022.csv')
summary(dat)
```


```{r}
#Convert categorical attributes into factor
dat[c('TYPE','HUNDRED_BLOCK','NEIGHBOURHOOD')] <- lapply(dat[c('TYPE','HUNDRED_BLOCK','NEIGHBOURHOOD')],as.factor)
dat <- dat[dat$NEIGHBOURHOOD!="",]
( unique(dat$TYPE) )
( as.vector(unique(dat$NEIGHBOURHOOD)))
summary(dat$TYPE)
```

```{r}
library(dplyr)
library(data.table)
pop <- read.csv('CensusLocalAreaProfiles2016.csv')
pop <- head(pop,5)
pop <- pop %>% select(-1,-2)
pop <- pop[4:5,]
pop <- transpose(pop)
colnames(pop) <- c('neighborhood','population')
pop[pop$neighborhood=="Arbutus-Ridge "]
pop
#Combine datasets

```
Population and taking sample
```{r}
N <- nrow(dat) #Population size
n <- 1000 # Sample size
set.seed(344)
library(sampling)
#Stratified sample by neighborhood
dat <- dat[order(dat$NEIGHBOURHOOD),]
dat <- dat[-(1:5),] #Drop first 5 rows where the neighborhood name is blank space
freq <- table(dat$NEIGHBOURHOOD)/nrow(dat)
freq <- as.vector(freq[-1])
n.h <- round(freq*n) #Each stratum sample size
strt <- strata(dat, stratanames = 'NEIGHBOURHOOD', size=n.h, method = 'srswr')
sample.strt <- dat[strt$ID_unit,]
```

```{r}
#Taking an SRS of size n=1000 from the population
SRS.index <- sample.int(N, n, replace = FALSE)
SRS <- dat[SRS.index,]
```
Now, we will estimate the proportion of crimes that happen during the summer months (July-August) using two samples above. We first use the SRS estimate and report both the estimated value as well as the standard error.
```{r}
summer <- SRS[SRS$MONTH %in% c(7,8),]
#Estimated proportion
p.hat.SRS <- nrow(summer)/n
#Standard error of the estimator, including FPC
se.SRS <- sqrt((1-n/N)*p.hat.SRS*(1-p.hat.SRS)/n)
( summer.SRS.results <- c(p.hat.SRS, se.SRS) )
```
Next, we will find the stratification estimator:
```{r}
sample.strt$summer <- ifelse(sample.strt$MONTH %in% c(7,8),1,0) #Create dummy for if the month is in summer
N.h <- dat %>% group_by(NEIGHBOURHOOD) %>% count() #population size for the strata

p.hat.h <- sample.strt %>%  #proportion for each strata
  group_by(NEIGHBOURHOOD) %>% 
  summarise(p.hat.h = mean(summer))
p.hat.str <- sum(N.h$n/N*p.hat.h$p.hat.h) #Estimated value
#Standard error
se.h <- sqrt((1 - n.h / N.h$n) * p.hat.h$p.hat.h*(1-p.hat.h$p.hat.h) / n.h)
se.str <- sqrt(sum((N.h$n / N)^2 * se.h^2))
( summer.str.results <- c(p.hat.str,se.str) )
```

We will now move on to the second parameter of interest, which is the proportion of Theft of Bicycle out of all crimes in 2022. We will use the same SRS and stratified sample as given above.

```{r}
#Estimator from SRS
bike_theft <- SRS[SRS$TYPE=='Theft of Bicycle',]
p.hat.SRS_bike <- nrow(bike_theft)/n
#Standard error of the estimator, including FPC
se.SRS_bike <- sqrt((1-n/N)*p.hat.SRS_bike*(1-p.hat.SRS_bike)/n)
( bike_theft.SRS.results <- c(p.hat.SRS_bike, se.SRS_bike) )
```
```{r}
#Estimator from stratified sample
sample.strt$bike_theft <- ifelse(sample.strt$TYPE=='Theft of Bicycle',1,0) #Create dummy for if the crime type is bike theft
p.hat.h_bike <- sample.strt %>%  #proportion for each strata
  group_by(NEIGHBOURHOOD) %>% 
  summarise(p.hat.h = mean(bike_theft))
p.hat.str_bike <- sum(N.h$n/N*p.hat.h_bike$p.hat.h) #Estimated value
#Standard error
se.h_bike <- sqrt((1 - n.h / N.h$n) * p.hat.h_bike$p.hat.h*(1-p.hat.h_bike$p.hat.h) / n.h)
se.str_bike <- sqrt(sum((N.h$n / N)^2 * se.h_bike^2))
( bike_theft.str.results <- c(p.hat.str_bike,se.str_bike) )
```






Proportion of crimes that happened during the summer (July-August)
Proportion of a type of crime out of all crime (Theft of Bicycle)
#Focus on 2 main params and look at population later if needed
SRS and stratification (proportional allocation)
Use FPC





